The Implementation of China’s Overseas NGO Law and the Operating Space for International Civil Society

### 2023-11-29 manuscript/manuscript.qmd
Overhaul of the structure in Introduction

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China. As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness INGOs.

The purpose of this study is to empirically examine how the ONGO Law has influenced the population of INGOs that have successfully registered in China since 2017, and in particular how the ONGO Law has shaped the operational space allowed for INGO operations in the country. By examining the *de facto* implementation of the law in addition to its formal *de jure* regulations, we provide a more comprehensive and data-based picture of the operational space for INGOs. By relying on representative empirical data from mainland China, our study also contributes to a growing body of research on the effects of closing civic space on INGO behavior and strategies.

This paper explores the characteristics of all registered INGOs in China in the five years after the ONGO Law became effective and examines the association between INGO issue areas and the operational space granted to them in the country. Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs vary in their ability to attract partner PSUs, and organizations that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy to measure the operational space for INGOs in China based on the assumption that INGOs need to or hope to have the approval to be able to carry out their programs in a geographical scope as large as possible [@Li:2020].

TODO: Andrew - make this a paragraph, not a list.

We make three general theoretical contributions in this paper:

1.  Expansion of work using institutional approaches to examine the regulation and restriction of civil society - [@DeMattee:2019]. To further understanding this global phenomena , China is an important case to explore the impacts such restrictive laws have on the operation of INGOs. Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case)

2.  Looking at how civil society laws act as a policy tool that regimes can leverage to shift NGO behavior: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly - The impacts on international civil society we examine include not only access or representation of INGOs in China, but also whether there is change of behaviors [@WuNie:2021]. Mention the concept of *flexibility (in the sense of adjusting to change)* too?

3.  Exploring differentiated application of these laws - *de facto* vs. *de jure* and differentiated treatment for more vs. less contentious INGOs. Authoritarian -\> arbitrary and varied application of the law -\> different *operating space*

@@

@@

# Closing civic space in China - legal background thing

China provides a fascinating example of both authoritarian preferences for INGO regulations and INGO adaptations to these regulations.

@@

>>>>

TODO: Andrew - review all the re-structuring made by Meng

In the past decade, nearly 100 countries have proposed or enacted legal measures designed to restrict civil society [@ICNL:2021], and only 21% of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @Civicus:2023]. This global trend of closing or shrinking civic space [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016] has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of legal restrictions on civil society with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China, or "ONGO Law," which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either register representative offices (ROs) or file temporary activities to lawfully operate in mainland China.

As an authoritarian state with a growing nonprofit sector, the Chinese model of implementing restrictive laws for international NGOs (INGOs) has important implications for understanding the dynamics of how civil society laws can be used as policy tools to harness the activities of INGOs. Globally, theoretical frameworks such as [@DeMattee:2019] focus on the 
institutional approaches, i.e., restrictive laws to crack down international civil society. (TO DO: whether to explain INGO vs. Civil Society as mentioned by MPSA reviewers).  To further understanding this global phenomena , China is an important case to explore the impacts such restrictive laws have on the operation of INGOs. Such legal restrictions can be either direct (denial of legal access/cannot register) or covert (red-tape, requiring the approval of PSUs, or the uncertainty of having to filing TAs case by case).

This paper explores the characteristics of all registered INGOs in China in the five years after the ONGO Law became effective and examines how the ONGO Law has influenced the population of INGOs that have successfully registered in China, as well as the association between factors such as INGO issue areas and the operational space granted to them in the country. Scholars have discussed the discrepancy that can exist between the *de facto* implementation of the law and the formal *de jure* regulations. (Andrew TO DO: add literature). We engage with this conversation by analyzing a more comprehensive and data-based picture of INGOs operating in China in the post ONGO law era. Exploring differentiated application of these laws and differentiated treatment for more vs. less contentious INGOs. Authoritarian -\> arbitrary and varied application of the law -\> different *operating space*

@@

Our study also contributes to a growing body of research on the effects of closing civic space on INGO behavior and strategies using a mixed-method approach.  Looking at how civil society laws act as a policy tool that regimes can leverage to shift NGO behavior: to tame INGO to be more cooperative to government. The requirement of PSU and Chinese partner to re-direct INGOs to be more authority friendly - The impacts on international civil society we examine include not only access or representation of INGOs in China, but also whether there is change of behaviors [@WuNie:2021]. In addition to quantitative analysis of the administrative data of 633 registered offices of INGOs. We also explored how the relationship with the government and the organizational flexibility of INGOs can come into play under the ONGO law.

@@

# The legal framework established by the ONGO law

@@

Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. INGOs vary in their ability to attract partner PSUs, and organizations that are unable to find corresponding PSUs are more limited in their geographic reach within the country. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. We use the number of provinces that INGOs are authorized to work in as a proxy to measure the operational space for INGOs in China based on the assumption that INGOs need to or hope to have the approval to be able to carry out their programs in a geographical scope as large as possible [@Li:2020].
